Appendix C: Latent Class Growth Analysis

### 2023-01-25 manuscript.Rmd
edits from Mauricio (with source editor)

Despite the popularity of LCA,
two challenges impede broader adoption and correct application of the method.
The first challenge is that most existing software that implements these models is commercial and closed-source.
The second challenge is that there is a lack of standards for estimating and reporting LCA, and some myths and misunderstandings have been perpetuated in prior literature.
This introduces a risk of misapplications of the technique,
and complicates manuscript review and assessment of the quality of published studies.
The present paper seeks to address these challenges by introducing updated guidelines for estimatin and reporting LCA,
based on current best practice. 
Furthermore, meeting the increasing demand for open source research software,
this paper introduces new functionality in the R-package `tidySEM` [@van_lissa_tidysem_2022] to perform LCA using free open source software (FOSS).

@@

Despite the popularity of LCA,
most existing software is commercial and closed-source,
for example, Mplus [@muthen_mplus_1998], for which the R-package `MplusAutomation` offers augmented LCA functionality [@hallquist_mplusautomation_2018], and LatentGOLD [REF].
<!-- software that implements these models. -->
<!-- One important limitation is that the most fully-features -->
A downside of commercial software is that its use is restricted to individuals and institutions who can afford a license.
Another downside is that the proprietary source code cannot be audited, debugged, or enhanced by third parties.
This goes against open science principles [@lamprecht_towards_2020],
incurs a risk that mistakes in the source code go unnoticed,
and curbs progress as independent developers cannot fix bugs or add functionality.

@@

Notable FOSS for LCA includes Mclust [REF], depmixS4 [REF], and OpenMx [@neale_openmx_2016].
A crucial limitation of these existing packages is that some have limited functionality (e.g., Mclust),
or implement specific innovations (e.g., depmixS4).
OpenMx is unique in that, like Mplus, it implement LCA in the context of a fully featured structural equation modeling framework.
Remaining obstacles to the use of `OpenMx` is that it is not very user-friendly and its functionality for LCA is poorly documented.

@@

When the focus is on each individual's estimated class membership,
LCA can be thought of as a type of clustering algorithm that corrects for (random) measurement error.
Specifically, the posterior class probability that an individual belongs to a latent class
can be computed from the likelihood of that individual's observed data under given the class-specific model.
These posterior class probabilities can be used to weight follow-up analyses,
or individuals can be assigned to classes based on their highest class probability.
If the classes are clearly distinct, measurement error is low, and the latter approach may be acceptable.
This perspective on mixture modeling is sometimes referred to as "model-based
clustering" [@hennig_handbook_2015; @scrucca_mclust_2016]. 
Many clustering algorithms apply some recursive splitting algorithm to the data. By
contrast "model-based" clustering refers to the fact that LCA estimates
cluster membership based on a parametric model.
Since a parametric model estimates only a fixed number of parameters,
it can be a relatively parsimonious solution compared to non-parametric techniques.

@@

LCA can also be used to reduce high-dimensional data
to just a few prototypes.
Factor analysis is also sometimes used for dimension reduction,
but only accounts for linear covariance between indicators,
whereas LCA can accommodate complex - but discrete - patterns.
For example, imagine that there exist two indicators - $x$ and $y$, and some individuals score high on both, some score low on both, and some score high on $x$ but low on $y$.
Factor analysis would only be able to capture the differences between the first two groups well; LCA captures differences between all three.
Note that this is a pragmatic use of LCA;
there is no theory about the existence of a categorical latent variable (cf. Testing theory).
Consequently, the class solution should be treated as descriptive of the sample,
and not be "reified" - treated as evidence of the existence of a categorical latent variable.

@@

Finally, LCA can be used to deal with data which violate certain assumptions.
For example, zero-inflated and hurdle models are special cases of LCA for dealing data with a very high proportion of zeros (extreme right skew) [@baughman_mixture_2007].
These models estimate one class to explain the "true zeros",
and another class with a different distribution to explain all oter values, which can include zero.
Note that this LCA has different models for both classes,
and thus falls beyond the scope of this paper.
<!-- CJ: This has been said elsewhere: -->
<!-- In fact, LCA assumes the population -->
<!-- distribution is a non-normal mixture of $K$ normal distributions, and it -->
<!-- can recover the value of $K$, i.e. generate a $K$-class solution. -->
<!-- CJ: This example is not perfect, we need another example still.. -->
For instance, the aforementioned study on marine plastic noted that many prior studies had treated the distribution of particles as (multivariate) normal, despite clear deviations from normality [@alkema_maximizing_2022].
A mixture of two normal distributions with distinct variances described the observed distribution better.
<!-- LCA and factor analysis share the approximation of an unobserved latent variable, the difference being if that variable is continuous or categorical. We encourage researchers to be agnostic about the **true** nature of the latent variable. And focus on the approach that fits the research question. -->
<!-- CJ: Why would we encourage researchers to be agnostic about the true...? -->

@@

Particular points of attention include measurement level of the indicators.
Recall that `tidySEM` can estimate mixture models with either numeric (type `numeric` and `integer`) or `ordered` indicators.
The column `type` should thus only contain these types.
Also check that the number of `unique` values match the intended measurement level.
Numeric variables with few unique values are typically better modeled as ordinal.
Our experience consulting on LCA methods
and moderating the `tidyLPA` Google group suggest that the
misapplication of mixture models to ordinal (e.g., Likert-type)
indicators is the most common source of user error. Whereas it has been
argued that some parametric methods are robust when scales with 7+
response categories are treated as continuous [e.g., @norman_likert_2010],
it is very unlikely
that an ordinal variable can be modeled as a *mixture* of multiple
normal distributions.
To illustrate the problem, consider what happens when estimating a 5-class solution on a single 5-point Likert scale indicator.
Each class-specific mean would describe a single response
category,
and a class-specific variance component would be empirically underidentified.
In sum, variables with few unique values might not be suitable for Gaussian mixture modeling;
LCA with ordinal indicators is more appropriate.
<!-- Descriptive statistics and -->
<!-- plots can be relegated to online supplements, provided that these are -->
<!-- readily accessible [consider using a GitHub repository as a -->
<!-- comprehensive public research archive, as explained in -->
<!-- @van_lissa_worcs_2021]. -->

@@

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
[@van_de_schoot_grolts-checklist_2017].
While we concur that investigating missingness is due diligence,
it is important to emphasize that missingness is adequately handled by default in `OpenMx`, which is the computational backend of `tidySEM`.
Its Full Information Maximum Likelihood (FIML) estimator makes use of all available
information without imputing missing values [@enders_relative_2001].
FIML is a best-practice solution for handling missing data;
at least on par with multiple imputation [@lee_comparison_2021].
Multiple imputation is less suitable to LCA for two reasons. First,
because LCA methods are often computationally intensive,
and conducting them on several imputed datasets multiplies the load.
Second, because many multiple imputation methods assume normality; an assumption likely to be violated when data come from a heterogeneous population.

Three types of missingness have
been distinguished in the literature [@rubin_inference_1976; 
@enders_applied_2022]: random missingness (MCAR);
missingness contingent on *observed* variables that can be known and controlled for (MAR);
and missingness related to unobserved variables,
which can be neither known nor controlled for (MNAR).
A so-called "MCAR" test is often reported [@jamshidian_tests_2010],
but note that this name is somewhat misleading,
as a significant test indicates that missingness is MAR,
but a non-significant test statistic does not distinguish between MCAR or MNAR.
As the classic MCAR test relies on the comparison of variances across groups with different patterns of missing data, it assumes normality [@little_test_1988].
This assumption is tenuous in the context of LCA.
A non-parametric MCAR test may be more suitable [@jamshidian_tests_2010].
For this tutorial,
the lead author has implemented this test in the `mice` package as `mice::mcar()`.
Our recommendation is to inspect and report the proportion of missingness per variable (e.g., using `tidySEM::descriptives()`),
and report the `mice::mcar()` test.
As FIML estimation assumes that missingness is either MCAR or MAR,
one would proceed with FIML regardless of the outcome.

@@

In principle, LCA can be based on any parametric model.
At present, `tidySEM` only facilitates estimation of models that can be specified in `lavaan` syntax.
The available model parameters depend on the measurement level of the indicators.
Continuous indicators can be parametrized in terms of means, variances, and covariances.
LCGA and GMM have these same parameters,
but with respect to the latent growth variables.
Ordinal indicators can be parametrized differently;
the default approach in `tidySEM` assumes that each ordinal indicator reflects an underlying standard normal distribution, $N(\mu = 0, \sigma = 1)$.
The parameters are "thresholds" that correspond to quantiles of this distribution, estimated based on the proportion of individuals in each of the
response categories of the indicator variable.
For example, a binary
indicator has a single threshold that distinguishes the two response
categories. If responses are distributed 50/50, then the corresponding
threshold would be $t_1 = 0.00$. If the responses are distributed 60/40,
then the resulting threshold would be $t_1 = 0.25$.
All of the aforementioned parameters can be freely estimated,
or constrained across classes, or fixed (e.g., to zero).

@@

As there are no absolute fit indices for LCA,
information criteria are typically used to assess relative fit.
Information criteria are computed from the $-2LL$,
which is the most basic fit measure.
Information criteria add a penalty for model complexity to the $-2LL$, and thus balance fit and complexity.
The lower
the value of an information criterion, the better the overall fit of the
model.
The original information criterion is the Akaike Information Criterion (AIC),
computed as $-2LL+2p$, where $p$ is the number of parameters.
The Bayesian Information Criterion (BIC) multiplies the penalty for complexity with a function of the sample size, $p \ln(n)$,
such that complex models are penalized more in smaller samples.
Finally,
the so-called *sample size adjusted* BIC (saBIC) implements a different ad-hoc penalty for sample size,
$\ln(\frac{n + 2}{24})$.
Although a simulation study showed that the saBIC performed well for class enumeration,
its penalty is not based on theory [REF].
BIC also performs well and is theoretically substantiated;
as such, it may be the most appropriate information criterion to use for model comparison [@nylund-gibson_ten_2018; @masyn_latent_2013]. 
All three ICs are available in the output of `table_fit()`.

Information criteria may occasionally contradict each other, so it is
important to identify a suitable strategy to reconcile them.
One option is to select a specific fit index, such as the BIC,
before seeing the results.
This choice can be preregistered to avoid concerns about cherry picking.
Another option is to always prefer the most parsimonious model that has best fit according to any of the available fit indices.
Yet another option is to incorporate information from multiple fit indices using the analytic hierarchy process [@akogul_comparison_2016],
which is implemented in `tidyLPA` [@rosenberg_tidylpa_2018].
Finally, one can use a scree plot to determine the inflection point at which additional classes contribute little to further decreases of the ICs [@nylund-gibson_ten_2018].
The preferred number of classes is at the inflection point where the slope of the curve is clearly leveling off.
Adding this many classes results in substantial decreases of the IC,
and adding further classes results in relatively small decreases.

@@

Anecdotal experience as a statistical co-author on LCA studies suggests that both these tests tend to be liberal,
returning significant p-values for models with more classes even when other indicators suggest that those models are overfitted (e.g., near-zero class counts).
Other authors have reported similar experiences [@sinha_practitioners_2021].
This suggests that, like existing ICs,
LR tests might not adequately balance fit and complexity in LCA.
Better solutions for class enumeration are thus an important target for future research.

@@

To assess classification accuracy, we turn to the third diagnostic table:
The average posterior probabilities by most likely class membership, $P_{p_{(M = m)}}$, `class_prob(res, type = "avg.mostlikely")`:
for participants assigned to one particular class (rows) according to $M$,
this table gives the mean probability of being assigned to any class (columns).
This table is obtained by calculating the column means of $P$ for the 
$k$ subsets of observations with most likely class of $M = 1 \ldots k$.
The diagonal represents the probability that observations in each class will be correctly classified.
<!-- Mauricio: Can you fact-check this?!? I really struglle here -->

The table of classification probabilities for the most likely latent class $M$ by latent class, $P_{p_{(M = m | C = c)}}$ is similar to the aforementioned table $P_{p_{(M = m)}}$,
except that it accounts for classification uncertainty.
If $C$ is the true class of an observation, which is imperfectly measured,
and $M$ is the most likely class,
then this table shows the probability of an observation being assigned to class $m$ in $M$, given that its true class is $c$ in $C$, or $p(M=m|C=c)$.
Values on the diagonal can thus be conceptualized as the reliability with which each level of the categorical latent variable is measured by $M$,
and off-diagonal elements can be conceptualized as measurement error.
<!-- Mauricio: Can you fact-check this?!? I really struglle here -->
The minimum and maximum of the diagonal of this table are given by default as part of the output of `table_fit()` for LCA analyses (`prob_min` and `prob_max`).
These probabilities are of particular interest because they conveniently summarize classification accuracy per class.
Both probabilities should be high, as this means that all classes are reliably classified.
If the minimum probability is low,
there is at least one class with poor classification accuracy.

@@

1. Report item descriptives
    + Are all items on similar scales?
    + Is the measurement level correctly coded?
    + Any other reasons why the observed (multivariate) distribution could not be described by the LCA (e.g., extreme skew)?
2. Report missingness
    + Proportion of missingness per variable
    + MAR test
    + Missing data method for handling missingness (typically FIML)
3. Report model specification
    + In exploratory LCA, compare all sensible alternative specifications
    + In confirmatory LCA, justify alternative models on theoretical grounds
    + Check that the observed (multivariate) distribution could theoretically be captured by the specified model (see #1).
4. Class enumeration
    + Justify the maximum number of classes $K$
    + In exploratory LCA, estimate 1:$K$ classes
    + In confirmatory LCA, estimate the theoretical number of classes, a 1-class model, and optionally other numbers of classes for comparison
5. Justify the (preregistered) criteria used for class enumeration, which can include:
    + Information criteria
    + LR tests
6. Justify the (preregistered) criteria used to eliminate models from consideration, including:
    + Classification diagnostics
    + Local identifiability (acceptable number of observations per parameter in each class)
    + Model convergence
    + Theoretical interpretability of the solution
7. Report the following:
    + Fit of all models under consideration
    + Classification diagnostics for all models under consideration
    + All parameters of the chosen model (means, (co)variances, thresholds)
    + Proportion of cases per class
8. Provide an informative plot of the chosen solution
    + Show raw data
    + Show parameters
    + Show confidence bounds if possible
    + Show multivariate distributions for multivariate models, if possible
9. Account for classification inaccuracy in follow-up analyses
    + If entropy > .95 and posterior classification probability for each class > .95, this is less crucial

>>>>

Despite the popularity of LCA,
two challenges impede broader adoption and correct application of the method.
The first challenge is that most existing software that implements these models is commercial and closed-source.
The second challenge is that there is a lack of standards for estimating and reporting LCA, and some myths and misunderstandings have been perpetuated in prior literature.
This introduces a risk of misapplications of the technique,
and complicates manuscript review and assessment of the quality of published studies.
The present paper seeks to address these challenges by introducing updated guidelines for estimating and reporting LCA,
based on current best practice. 
Furthermore, meeting the increasing demand for open source research software,
this paper introduces new functionality in the R-package `tidySEM` [@van_lissa_tidysem_2022] to perform LCA using free open source software (FOSS).

@@

Despite the popularity of LCA,
most existing software is commercial and closed-source,
for example, Mplus [@muthen_mplus_1998], for which the R-package `MplusAutomation` offers augmented LCA functionality [@hallquist_mplusautomation_2018], and LatentGOLD [@LG_manual].
<!-- software that implements these models. -->
<!-- One important limitation is that the most fully-features -->
A downside of commercial software is that its use is restricted to individuals and institutions who can afford a license.
Another downside is that the proprietary source code cannot be audited, debugged, or enhanced by third parties.
This goes against open science principles [@lamprecht_towards_2020],
incurs a risk that mistakes in the source code go unnoticed,
and curbs progress as independent developers cannot fix bugs or add functionality.

@@

Notable FOSS for LCA includes mclust [@scrucca_mclust_2016], depmixS4 [@depmixS4], and OpenMx [@neale_openmx_2016].
A crucial limitation of these existing packages is that some have limited functionality (e.g., Mclust),
or implement specific innovations (e.g., depmixS4).
OpenMx is unique in that, like Mplus, it implement LCA in the context of a fully featured structural equation modeling framework.
Remaining obstacles to the use of `OpenMx` is that it is not very user-friendly and its functionality for LCA is poorly documented.

@@

When the focus is on each individual's estimated class membership,
LCA can be thought of as a type of clustering algorithm that corrects for random measurement error by default, and is able to correct for systematic measurement error if it is included in the model.
Specifically, the posterior class probability that an individual belongs to a latent class
can be computed from the likelihood of that individual's observed data under given the class-specific model.
These posterior class probabilities can be used to weight follow-up analyses,
or individuals can be assigned to classes based on their highest class probability.
If the classes are clearly distinct, measurement error is low, and the latter approach may be acceptable.
This perspective on mixture modeling is sometimes referred to as "model-based
clustering" [@hennig_handbook_2015; @scrucca_mclust_2016]. 
Many clustering algorithms apply some recursive splitting algorithm to the data. By
contrast "model-based" clustering refers to the fact that LCA estimates
cluster membership based on a parametric model.
Since a parametric model estimates only a fixed number of parameters,
it can be a relatively parsimonious solution compared to non-parametric techniques.

@@

LCA can also be used to reduce high-dimensional data
to just a few prototypes.
Factor analysis is also sometimes used for dimension reduction,
but only accounts for linear covariance between indicators,
whereas LCA can accommodate complex - but discrete - patterns.
For example, imagine that there exist two indicators - $x$ and $y$, and some individuals score high on both, some score low on both, and some score high on $x$ but low on $y$.
Factor analysis would only be able to capture the differences between the first two groups well; LCA captures differences between all three.
Note that this is a pragmatic use of LCA;
there is no theory about the existence of a categorical latent variable (cf. Testing theory).
Consequently, the class solution should be treated as descriptive of the sample,
and not be "reified" - treated as evidence of the existence of a categorical latent variable. These classes can be theoretically justified, and use as part of the theory building process, but they are not sufficient.

@@

Finally, LCA can be used to deal with data which violate certain assumptions.
For example, zero-inflated and hurdle models are special cases of LCA for dealing data with a very high proportion of zeros (extreme right skew) [@baughman_mixture_2007; magnus_multidimensional_2021].
These models estimate one class to explain the "true zeros",
and another class with a different distribution to explain all other values, which can include zero.
Note that this LCA has different models for both classes,
and thus falls beyond the scope of this paper.
<!-- CJ: This has been said elsewhere: -->
<!-- In fact, LCA assumes the population -->
<!-- distribution is a non-normal mixture of $K$ normal distributions, and it -->
<!-- can recover the value of $K$, i.e. generate a $K$-class solution. -->
<!-- CJ: This example is not perfect, we need another example still.. -->
For instance, the aforementioned study on marine plastic noted that many prior studies had treated the distribution of particles as (multivariate) normal, despite clear deviations from normality [@alkema_maximizing_2022].
A mixture of two normal distributions with distinct variances described the observed distribution better.
<!-- LCA and factor analysis share the approximation of an unobserved latent variable, the difference being if that variable is continuous or categorical. We encourage researchers to be agnostic about the **true** nature of the latent variable. And focus on the approach that fits the research question. -->
<!-- CJ: Why would we encourage researchers to be agnostic about the true...? -->

@@

Particular points of attention include measurement level of the indicators.
Recall that `tidySEM` can estimate mixture models with either numeric (type `numeric` and `integer`) or `ordered` indicators.
The column `type` should thus only contain these types.
Also check that the number of `unique` values match the intended measurement level.
Numeric variables with few unique values are typically better modeled as ordinal.
Our experience consulting on LCA methods
and moderating the `tidyLPA` Google group suggest that the
misapplication of mixture models to ordinal (e.g., Likert-type)
indicators is the most common source of user error. Whereas it has been
argued that some parametric methods are robust when scales with 7+
response categories are treated as continuous [e.g., @norman_likert_2010; rhemtulla_when_2012],
it is very unlikely
that an ordinal variable can be modeled as a *mixture* of multiple
normal distributions.
To illustrate the problem, consider what happens when estimating a 5-class solution on a single 5-point Likert scale indicator.
Each class-specific mean would describe a single response
category,
and a class-specific variance component would be empirically underidentified.
In sum, variables with few unique values might not be suitable for Gaussian mixture modeling;
LCA with ordinal indicators is more appropriate.
<!-- Descriptive statistics and -->
<!-- plots can be relegated to online supplements, provided that these are -->
<!-- readily accessible [consider using a GitHub repository as a -->
<!-- comprehensive public research archive, as explained in -->
<!-- @van_lissa_worcs_2021]. -->

@@

Previous work has emphasized the importance of examining the pattern of
missing data and reporting how missingness was handled
[@van_de_schoot_grolts-checklist_2017].
While we concur that investigating missingness is due diligence,
it is important to emphasize that missingness is adequately handled by default in `OpenMx`, which is the computational backend of `tidySEM`.
Its Full Information Maximum Likelihood (FIML) estimator makes use of all available
information without imputing missing values, by estimating the model based on all
available information for each subject [@enders_relative_2001; enders_applied_2022].
FIML is a best-practice solution for handling missing data;
at least on par with multiple imputation [@lee_comparison_2021].
Multiple imputation is less suitable to LCA for two reasons. First,
because LCA methods are often computationally intensive,
and conducting them on several imputed datasets multiplies the load.
Second, because many multiple imputation methods assume normality; an assumption likely to be violated when data come from a heterogeneous population, so the imputation model could bias the results as it does not account for the expected heterogeneity.

Three types of missingness have
been distinguished in the literature [@rubin_inference_1976; 
@enders_applied_2022]: random missingness (MCAR);
missingness contingent on *observed* variables that can be known and controlled for (MAR);
and missingness related to unobserved variables,
which can be neither known nor controlled for (MNAR).
A so-called "MCAR" test is often reported [@jamshidian_tests_2010],
but note that this name is somewhat misleading,
as a significant test indicates that missingness is MAR,
but a non-significant test statistic does not distinguish between MCAR or MNAR.
As the classic MCAR test relies on the comparison of variances across groups with different patterns of missing data, it assumes normality [@little_test_1988].
This assumption is tenuous in the context of LCA.
A non-parametric MCAR test may be more suitable [@jamshidian_tests_2010].
For this tutorial,
the lead author has implemented this test in the `mice` package as `mice::mcar()`.
Our recommendation is to inspect and report the proportion of missingness per variable (e.g., using `tidySEM::descriptives()`),
and report the `mice::mcar()` test, as way to identify variables that can help the parameter recovery instead of evidence of MCAR or MNAR.
As FIML estimation assumes that missingness is either MCAR or MAR,
one would proceed with FIML regardless of the outcome.

@@

In principle, LCA can be based on any parametric model.
At present, `tidySEM` only facilitates estimation of models that can be specified in `lavaan` syntax.
The available model parameters depend on the measurement level of the indicators.
Continuous indicators can be parametrized in terms of means, variances, and covariances.
LCGA and GMM have these same parameters,
but with respect to the latent growth variables.
Ordinal indicators are parametrized differently;
the default approach in `tidySEM` assumes that each ordinal indicator reflects an underlying standard normal distribution, $N(\mu = 0, \sigma = 1)$.
The parameters are "thresholds" that correspond to quantiles of this distribution, estimated based on the proportion of individuals in each of the
response categories of the indicator variable.
For example, a binary
indicator has a single threshold that distinguishes the two response
categories. If responses are distributed 50/50, then the corresponding
threshold would be $t_1 = 0.00$. If the responses are distributed 60/40,
then the resulting threshold would be $t_1 = 0.25$.
All of the aforementioned parameters can be freely estimated,
or constrained across classes, or fixed (e.g., to zero).

@@

As there are no absolute fit indices for LCA,
information criteria are typically used to assess relative fit.
Information criteria are computed from the $-2LL$,
which is the most basic fit measure, testing the overall hypothesis if the model properly represents the data.
Information criteria add a penalty for model complexity to the $-2LL$, and thus balance fit and complexity.
The lower
the value of an information criterion, the better the overall fit of the
model.
The original information criterion is the Akaike Information Criterion (AIC),
computed as $-2LL+2p$, where $p$ is the number of parameters.
The Bayesian Information Criterion (BIC) multiplies the penalty for complexity with a function of the sample size, $p \ln(n)$,
such that complex models are penalized more in smaller samples.
Finally,
the so-called *sample size adjusted* BIC (saBIC) implements a different ad-hoc penalty for sample size,
$\ln(\frac{n + 2}{24})$.
Although a simulation study showed that the saBIC performed well for class enumeration,
its penalty is not based on theory [REF].
BIC also performs well and is theoretically substantiated;
as such, it may be the most appropriate information criterion to use for model comparison [@nylund-gibson_ten_2018; @masyn_latent_2013]. 
All three ICs are available in the output of `table_fit()`.

Information criteria may occasionally contradict each other, so it is
important to identify a suitable strategy to reconcile them.
One option is to select a specific fit index, such as the BIC,
before seeing the results.
This choice can be preregistered to avoid concerns about cherry picking.
Another option is to always prefer the most parsimonious model that has best fit according to any of the available fit indices. Other approach is to select several competing model based on the ICs, and evaluate the theoretical interpretation.
Yet another option is to incorporate information from multiple fit indices using the analytic hierarchy process [@akogul_comparison_2016],
which is implemented in `tidyLPA` [@rosenberg_tidylpa_2018].
Finally, one can use a scree plot to determine the inflection point at which additional classes contribute little to further decreases of the ICs [@nylund-gibson_ten_2018].
The preferred number of classes is at the inflection point where the slope of the curve is clearly leveling off.
Adding this many classes results in substantial decreases of the IC,
and adding further classes results in relatively small decreases.

@@

Anecdotal experience as a statistical co-author on LCA studies suggests that both these tests tend to be liberal,
returning significant p-values for models with more classes even when other indicators suggest that those models are overfitted (e.g., near-zero class counts).
Other authors have reported similar experiences [@sinha_practitioners_2021], which matches the differences between LRT and ICs from information theory as LRT tests and compare the models based on in sample predictive accuracy, and ICs evaluates and compare the models out-of-sample predictive accuracy which helps to control for overfitting [@mcelreath_statistical_2020].
This suggests that, like existing ICs,
LR tests might not adequately balance fit and complexity in LCA.
Better solutions for class enumeration are thus an important target for future research.

@@

To assess classification accuracy, we turn to the third diagnostic table:
The average posterior probabilities by most likely class membership, $P_{p_{(M = m)}}$, `class_prob(res, type = "avg.mostlikely")`:
for participants assigned to one particular class (rows) according to $M$,
this table gives the mean probability of being assigned to any class (columns).
This table is obtained by calculating the column means of $P$ for the 
$k$ subsets of observations with most likely class of $M = 1 \ldots k$.
The diagonal represents the average predictive probability of a the subjects related to their most likely class, indicating the certainty of class assignment for the subjects. Some guidelines have being suggested for this diagonal, like for them to be higher than 0.7 [@masyn2013], but we suggest researchers to be cautions about these guidelines as they cannot be generalized and should not be use as strict cutoffs as evidence "good" or "bad" classification.

The table of classification probabilities for the most likely latent class $M$ by latent class, $P_{p_{(M = m | C = c)}}$ is similar to the aforementioned table $P_{p_{(M = m)}}$,
except that it accounts for classification uncertainty.
If $C$ is the true class of an observation, which is imperfectly measured,
and $M$ is the most likely class,
then this table shows the probability of an observation being assigned to class $m$ in $M$, given that its true class is $c$ in $C$, or $p(M=m|C=c)$.
Values on the diagonal can thus be conceptualized as the reliability with which each level of the categorical latent variable is measured by $M$,
and off-diagonal elements can be conceptualized as measurement error.
<!-- Mauricio: Can you fact-check this?!? I really struglle here. MGV: this looks good -->
The minimum and maximum of the diagonal of this table are given by default as part of the output of `table_fit()` for LCA analyses (`prob_min` and `prob_max`).
These probabilities are of particular interest because they conveniently summarize classification accuracy per class.
Both probabilities should be high, as this means that all classes are reliably classified.
If the minimum probability is low,
there is at least one class with poor classification accuracy.

@@

1. Report item descriptives
    + Are all items on similar scales?
    + Is the measurement level correctly coded?
    + Any other reasons why the observed (multivariate) distribution could not be described by the LCA (e.g., extreme skew)?
2. Report missingness
    + Proportion of missingness per variable
    + MAR test
    + Missing data method for handling missingness (typically FIML)
3. Report model specification
    + In exploratory LCA, compare all sensible alternative specifications
    + In confirmatory LCA, justify alternative models on theoretical grounds
    + Check that the observed (multivariate) distribution could theoretically be captured by the specified model (see #1).
4. Class enumeration
    + Justify the maximum number of classes $K$
    + In exploratory LCA, estimate 1:$K$ classes
    + In confirmatory LCA, estimate the theoretical number of classes, a 1-class model, and optionally other numbers of classes for comparison
5. Justify the (preregistered) criteria used for class enumeration, which can include:
    + Information criteria
    + LR tests
    + Theoretical interpretability of the solution
6. Justify the (preregistered) criteria used to eliminate models from consideration, including:
    + Classification diagnostics
    + Local identifiability (acceptable number of observations per parameter in each class)
    + Model convergence
    + Theoretical interpretability of the solution
7. Report the following:
    + Fit of all models under consideration
    + Classification diagnostics for all models under consideration
    + All parameters of the chosen model (means, (co)variances, thresholds)
    + Proportion of cases per class
8. Provide an informative plot of the chosen solution
    + Show raw data
    + Show parameters
    + Show confidence bounds if possible
    + Show multivariate distributions for multivariate models, if possible
9. Account for classification inaccuracy in follow-up analyses
    + If entropy > .95 and posterior classification probability for each class > .95, this is less crucial
